### 2024-11-10 doc/sysdesc.tex
more

\documentclass{llncs}
\usepackage[show]{ed}
\usepackage{cleveref}
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{stex-logo}
\usepackage{listings}
\definecolor{codegray}{rgb}{0.9,0.9,0.9}
\lstset{basicstyle=\sf,columns=fullflexible,backgroundcolor = \color{codegray}}

\usepackage[style=alphabetic,hyperref=auto,defernumbers=true,backend=bibtex,firstinits=true,maxbibnames=9,maxcitenames=3,isbn=false]{biblatex}
\addbibresource{kwarcpubs.bib}
\addbibresource{extpubs.bib}
\addbibresource{kwarccrossrefs.bib}
\addbibresource{extcrossrefs.bib}

% I do not want to annotate just yet.

\newcommand\ALeA{\textsf{ALeA}\xspace}
\newcommand\srf{\textsf{srify}\xspace}
\def\llangle{\langle\kern-.2em\langle}
\def\rrangle{\rangle\kern-.2em\rangle}

\title{Bulk Semantic Annotation with an Partially Known Knowledge Base}
\author{Michael Kohlhase, Jan Frederik Schaefer}
\institute{Computer Science, FAU Erlangen N\"urnberg, Germany}
\begin{document}
\maketitle
\begin{abstract}
tbw
\end{abstract}

\section{Introduction}
Arguably, dealing with large document collections is one of the key factors in the
knowledge-driven society and economy. There are currently two main contenders for machine
support in this area. Symbolic/logic-based technologies and sub-symbolic,
machine-learning-based AI, e.g. via LLMs or chatbots; they have complementary strengths
and challenges: Symbolic technologies offer precision and explainability out of the box,
but face scalability challenges, because the prerequisite background knowledge has to be
(manually) formalized. ML-based approaches can be trained on all data of the Internet, but
face challenges in precision and explanations. Aarne Ranta matches these profiles to the
notion of \textbf{producer tasks} -- i.e. tasks where precision is key, but limited
coverage ($10^{3\pm1}$ concepts) is OK; e.g. for multi-language/variant manuals for very
expensive tool machines -- and \textbf{consumer tasks}, where coverage is key and
precision secondary; e.g. for consumer-grade machine translation like \textsf{Google
  translate} in \cite{Ranta:atcp17}.

In this paper, we address tools that help attack the coverage problem in symbolic/semantic
approaches in practice. The producer task we use as a case study is that of adaptive
learning assistants for MINT subjects, concretely the \ALeA system
\cite{BerBetChu:lssmkm23}, which uses \sTeX \cite{MueKo:sdstex22,sTeX:github:on} -- a
variant of {\LaTeX} that allows to embed semantic annotations -- for a knowledge
representation and generates learner-adaptive learning objects instrumented with
learning-support interactions from that. The \textbf{\sTeX/\ALeA content commons}
\ednote{MK@MK: continue to describe the size and distribution, and
  authorship}\ednote{explain the \textbf{domain model} (the SMGloM) and the
  \textbf{formulation model} (cf. \cite{BerBetChu:lssmkm23}) in the commons and give their
  releative sizes.}

We contend that while education -- especially higher education -- as a whole is a
wide-coverage task for society education of individual learners in specific domains --
especially, if we want to tailor the educational offerings to said individual/domain -- is
a producer task, at least until ML-based methods reach the precision and explainability
required by the ethics of teaching.

\section{Semantic Authoring in a Content Common}

The main practical problem of annotating presentational {\LaTeX} course materials into
semantic documents that carry enough information to support meaningful learning support
services is to add semantic references to technical terms from the domain jargon. Take for
instance a learning object like the paragraph in \cref{fig:lo}, where we want to annotate
the the term ``state space'' with the (semantic) concept of a ``state space of a search
problem''. The latter can either be defined earlier in the course or in the domain mode
(or both).

\begin{figure}[ht]\centering
  \includegraphics[width=8cm]{problem}
  \caption{An annotated Learning Object from an AI Lecture}\label{fig:lo}
\end{figure}

Concretely this is about converting the {\LaTeX} string
\begin{lstlisting}[language={[LaTeX]TeX}]
\item \blue{Problem:} The state space has local minima, e.g. the board on the
  right has $h=1$ but every successor has $h>1$. 
\end{lstlisting}
into the (partially annotated item:  
\begin{lstlisting}[language={[LaTeX]TeX},morekeywords={sn,importmodule}]
\importmodule[smglom/search]{mod?state-space}
[...]  
\item \blue{Problem:} The \sn{state space} has local minima, e.g. the
  board on the right has $h=1$ but every successor has $h>1$. 
\end{lstlisting}
Concretely, the author has to remember the module for the term ``state space'' from
\cref{fig:state-space}, the symbol name \lstinline|state space|, annotate it with the
\sTeX macro \lstinline|\sn{...}|, and add the module import in the first line (unless it
is already imported). In this case the symbol name happens to be the same as its English
verbalization, otherwise we would have to use
\lstinline[mathescape]|\sr{$\llangle symbol name\rrangle$}{state space}|.

\begin{figure}[ht]\centering
  \includegraphics[width=12cm]{state-space}
  \caption{The definition of the term ``state space'' from the domain model}\label{fig:state-space}
\end{figure}

For this, the author has to be aware that the module from \cref{fig:state-space} exists,
the symbol name, know the URL and manage reduncancy of the imports. At ca 150.000 words in
a semester's words of lecture notes and ca. 10-15\% of technical terms a rather daunting
task.

\sTeX authoring is already supported by an VSCode IDE plugin \cite{sTeX-IDE:git} that
analyzes annotations on the fly, displays the underlying reference semantics, reports
errors and redundancies, and even offers a concept search interface.

\section{Conclusion}

In this system description we have presented the \srf tool, a command-line tool

\printbibliography
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

% LocalWords:  tbw Aarne Ranta