### 2025-07-03 doc/cicm25-paper/rebuttal.txt
preamble

We thank the reviewers for their thorough and thoughtful comments and suggestions.

We concur with the reviewers that the paper is somewhere between a research paper and a
system description, and we have struggled with this in the paper writing process. Indeed
we had started out as the former, and then realized that there is more to the topic than a
system description could convey. We did not try to bloat the text to make it more suitable
to the category, but tried to layout the task of semantic authoring -- which we see as a
novel category -- as a general problem to be addressed by the CICM community.