### 2025-07-03 doc/cicm25-paper/rebuttal.txt
more

We thank the reviewers for their thorough and thoughtful comments and suggestions.

We concur with the reviewers that the paper is somewhere between a research paper and a
system description, and we have struggled with this in the paper writing process. Indeed
we had started out with the former, and then realized that there is more to the topic than a
system description could convey.

Howver, we did not try to bloat the text to make it more suitable to the category, but
tried to layout the task of semantic authoring -- which we see as a novel task category --
as a general problem to be addressed by the CICM community.  Given reviewer 1's feedback,
we will take hard look at the paper and see where we can shorten/streamline it.  While
writing, we assumed that not all readers would find the ideas as easy as reviewer 1
expressed, considering the broad range of backgrounds at CICM.

We will, of course, incorporate the minor corrections suggested by the reviewers; these
are very helpful to us.

To address a few other questions/concerns:

> While the whole project is a significant effort, the special tools described here  'snify' and 'defianno' seem to this reviewer not particularly able to scale up in the way desired. The measurement proposed for the increase of productivity seems very much subjective and the 'STEX/ALeA content commons' also seem specialized and not very `common', i.e. not shared by many. The problem with ambiguity of definitions and their contexts also seems not very easy to address with the tools proposed.

TODO: This summarizes the criticism of rev. 2... Not sure what to do with the first and last sentence.

> 'STEX/ALeA content commons' also seem specialized and not very `common', i.e. not shared by many.

True.
Most content is about mathematics and computer science,
but some other disciplines are represented to a lesser extent (e.g. law and linguistics).
Calling it a 'commons' is somewhat aspirational.

> be more precise and quantitative when you write "one order of magnitude more efficient"

We agree.
The efficiency gain cannot easily be quantified and we have therefore kept the assertions
about that intentionally vague.
It strongly depends on the author's familiarity with the domain model (having to search
for the system URI of a concept that "should be in the domain model" is so time-consuming
that we get the efficiency gain of "one order of magnitude in this part alone") .
Also, authors tend to skip more technical terms during annotation if they are not using
snify; so they retain efficiency at the cost of lowering the quality of annotation.
We measured values 
ranging from a factor of 3 (for an author who knows the domain model by heart),
to a factor of 20 (for an author who has never annotated before).
Given the reviewer questions we will elaborate on this in the paper.

> Why do symbolic technologies offer explainability "out of the box"?

This statement has to be read in the context of comparing with machine-learning or neural
AI technologies, where the "context and inference" only exist in the form of a
weight/parameter distribution over the network. In symbolic technolgies, the inferences
(and/or transformation steps) can be traced and turned into argumentations or
justifications. While this may not be technically be "out of the box" at least the
necessary resources are all there.

> where does [the 150,000 words] estimate come from?

This is an estimate based on our own lecture materials (a combination of slides and
notes), specifically two lectures on artificial intelligence.
The concrete numbers will of course vary by course and by instructor,
but we were trying to give an idea of the scale.
We will clarify this in the paper.

> Is there more related work?

XXXX not to our knowledge, otherwise we would have discussed it XXXX