### 2025-07-04 doc/cicm25-paper/rebuttal.txt
a few more changes

We thank the reviewers for their thorough and thoughtful comments and
suggestions.

We concur with the reviewers that the paper is somewhere between a research
paper and a system description, and we have struggled with this in the paper
writing process. Indeed we had started out with the former, and then realized
that there is more to the topic than a system description could convey.
However, we did not try to bloat the text to make it more suitable to the
category, but tried to layout the task of semantic authoring -- which we see as
a novel task category -- as a general problem to be addressed by the CICM
community.  Given reviewer 1's feedback, we will take hard look at the paper
and see where we can shorten/streamline it.  While writing, we assumed that not
all readers would find the ideas as easy as reviewer 1 indicated, considering
the broad range of backgrounds at CICM.

Reviewer 2 is skeptical regarding the scalability. snify is by now the main
tool for term annotation in sTeX and has been used for thousands (likely over
10 000) annotations. Nevertheless, maintaining a catalog across the growing
ecosystem is a challenge we are currently working on. It is true that the
'sTeX/ALeA content commons' is mostly shared within our group (though across
many different lectures), but we hope to grow the community and are in the
early stages of collaboration with lecturers from IISc and Bath University.

Reviewer 2 also mentions the problem of ambiguity. The annotations disambiguate
the meaning of terms with the granularity of the domain model, which may be
insufficient in some cases (e.g. if there is only one entry for natural
numbers, regardless of whether they include 0 or not).

Reviewer 3 rightfully points out that it would be better to state the
efficiency gains more precisely. Unfortunately, the efficiency gain cannot
easily be quantified in a number and have therefore kept the assertions vague.
It strongly depends on the author's familiarity with the domain model (having
to search for the system URI of a concept that "should be in the domain model"
is very time-consuming). Also, authors tend to skip more technical terms
during annotation if they are not using snify; so they retain efficiency at the
cost of lowering the quality of annotation. We measured values ranging from a
factor of 3 (for an author who knows the domain model by heart), to a factor of
20 (for an author who has never annotated before). The focus mode of snify
further improves the efficiency in terms of annotations per time. Given the
reviewer questions we will elaborate on this in the paper.

We will, of course, incorporate the minor corrections suggested by the
reviewers; these are very helpful to us.

To address some of the minor questions/concerns:

> Reviewer 1 asks why symbolic technologies offer explainability "out of the
> box"

This statement has to be read in the context of comparing with machine-learning
or neural AI technologies, where the "context and inference" only exist in the
form of a weight/parameter distribution over the network. In symbolic
technologies, the inferences (and/or transformation steps) can be traced and
turned into argumentations or justifications. While this may not be technically
be "out of the box" at least the necessary resources are all there. We will
improve the formulation.

> Reviewer 2 asks where the 150,000 words/10-15% estimate comes from

This is an estimate based on our own lecture materials (a combination of slides
and notes), specifically our (admittedly rather large) lecture on artificial
intelligence. The concrete numbers will of course vary by course and by
instructor, but we were trying to give an idea of the scale. We will clarify
this in the paper.

> Reviewer 2 asks if there is more related work

To our knowledge, we have covered the spectrum of related work, but we would be
grateful for pointers to any work we may have missed.

> Reviewer 2 suggests that our use of "verbalization" should be explained

For us, the verbalization of a symbol is its rendering in (written) natural
language. For example, "integer" and "whole number" are English verbalizations
of the symbol for integers. We will clarify this in the paper.