### 2025-07-05 doc/cicm25-paper/rebuttal.txt
final version

We thank the reviewers for their thorough and thoughtful comments and
suggestions.

We admit that the paper is somewhere between a research paper and a system
description, and we have struggled with this in the paper writing process.
Indeed we had started out with the former, and then realized that there is more
to the topic than a system description could convey. However, we did not try to
bloat the text to make it more suitable to the category, but tried to layout
the task of semantic authoring -- which we see as a novel task category -- as a
general problem to be addressed by the CICM community.  Given reviewer 1's
feedback, we will take hard look at the paper and see where we can
shorten/streamline it.  While writing, we assumed that not all readers would
find the ideas as easy as reviewer 1 indicated, considering the broad range of
backgrounds at CICM.

Reviewer 2 is skeptical regarding the scalability. snify is by now the main
tool for term annotation in sTeX and has been used for thousands (likely over
10 000) annotations. Nevertheless, maintaining a catalog across the growing
ecosystem is a challenge we are currently working on. It is true that the
'sTeX/ALeA content commons' is mostly shared within our group (though across
many different lectures), but we hope to grow the community and are in the
early stages of collaboration with lecturers from IISc and Bath University.

Reviewer 3 rightfully points out that it would be better to state the
efficiency gains more precisely. Unfortunately, the efficiency gain cannot
easily be quantified in a number and we have therefore kept the assertions
vague. It strongly depends on the author's familiarity with the domain model
(having to search for the system URI of a concept that "should be in the domain
model" is very time-consuming). Also, authors tend to skip more technical terms
during annotation if they are not using snify; so they retain efficiency at the
cost of lowering the quality of annotation. We measured values ranging from a
factor of 3 (for an IDE-supported expert who knows the domain model by heart),
to a factor of 13 (for an inexperienced annotator with IDE support). The focus
mode of snify further improves the efficiency in terms of annotations per time.
Given the reviewer questions we will elaborate on this in the paper.

We will, of course, incorporate the minor corrections suggested by the
reviewers; these are very helpful to us.

To address some of the minor questions/concerns:

> Reviewer 1 asks why symbolic technologies offer explainability "out of the
> box"

This statement has to be read in the context of comparing with machine-learning
or neural AI technologies, where the "context and inference" only exist in the
form of a weight/parameter distribution over the network. In symbolic
technologies, the inferences (and/or transformation steps) can be traced and
turned into argumentations or justifications. While this may not be technically
be "out of the box" at least the necessary resources are all there. We will
improve the formulation in the paper.

> Reviewer 2 asks where the 150,000 words/10-15% estimate comes from

This is an estimate based on our own lecture materials (a combination of slides
and notes), specifically our (admittedly rather large) lecture on artificial
intelligence. The concrete numbers will of course vary by course and by
instructor, but we were trying to give an idea of the scale. We will clarify
this in the paper.

> Reviewer 2 asks if there is more related work

To our knowledge, we have covered the spectrum of related work, but we would be
grateful for pointers to any work we may have missed.

> Reviewer 2 suggests that our use of "verbalization" should be explained

For us, the verbalization of a symbol is its rendering in (written) natural
language. For example, "integer" and "whole number" are English verbalizations
of the symbol for integers. We will clarify this in the paper.