### 2025-09-26 doc/aligments2025/doc.tex
alignments abstract draft

\documentclass{ceurart}
\usepackage[show]{ed}
\usepackage{paralist}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
}

\title{Efficiently Linking Documents to Alignment Databases -- A Prototype-Based Exploration}

\author[1]{Michael Kohlhase}
\author[2]{Jan Frederik Schaefer}

\conference{Workshop on Math Datasets: alignments and comparisons (ALIGN2025),
  October 10, 2025, Brasilia, Brazil}

\begin{document}
\maketitle
Both formal and informal mathematical knowledge is fragmented across many different systems and formats.
A first step towards combining these resources is aligning their concepts.

The focus of this paper is the alignment
of concept references in documents -- e.g.\ scientific papers -- with a concept database.
We will use WikiData as the concept database in our examples.
As WikiData links to other resources (e.g.\ nLab, ProofWiki, OEIS, \ldots),
and is also referenced in other databases
(e.g. the MathGloss~\cite{mathgloss:arxiv} alignment database),
% As WikiData is often referenced in other alignment databases,
% e.g.\ MathGloss~\cite{mathgloss:arxiv},
we also get indirect alignments with other systems.

As a running example, we will use the simple declarational sentence
\begin{quote}
Let $n$ be a natural number.
\end{quote}
where we want to align the concept reference ``natural number'' with
the corresponding WikiData identifier.
Such an alignment could, for example,
provide additional information to the reader,
could determine the prerequisite knowledge for understanding the document,
or be a first step towards formalization.
Creating such alignments is difficult.
For our example there are, in fact, three possible alignment targets:
\begin{compactitem}
\item \texttt{Q21199} (natural numbers, possibly including 0)
\item \texttt{Q28920044} (positive integers, i.e.\ natural numbers excluding 0)
\item \texttt{Q28920052} (non-negative integers, i.e.\ natural numbers including 0)
\end{compactitem}
We claim that machine learning techniques are not yet capable of
making the right choice sufficiently reliably,
which puts the burden of creating alignments on human annotators --
ideally the authors.

The most obvious workflow is rather inefficient:
the author has to either recall the correct alignment target from
memory or search for it, and then edit the document to add the alignment annotation.
A more efficient workflow, supported by dedicated tools,
is crucial for large-scale alignment efforts.

The sTeX ecosystem, which uses diverse semantic markup to drive
learning support systems,
faced a similar challenge, which was recently addressed by the snify system.
In this paper, we present a prototypical extension of the snify system
for efficient alignment that is independent of the sTeX ecosystem.
Concretely, we have added the option to use WikiData as the concept database,
rudimentary support for annotating formulae,
the ability to annotate HTML documents,
and a browser-based user interface, which
helps with the annotation of HTML documents.

The snify tool is a simple command line application that
steps through the phrases that could be annotated in a document
and presents the user with a list of potential targets.
\Cref{fig:snify} shows the tool in action.
The user can select the correct entry from the list by typing its number,
which then creates the alignment by inserting a macro into the document
(here e.g.\ \lstinline.\wdalign{Q28920044}{natural number}).
Aside from choosing the correct annotation,
snify has many other commands to skip
phrases, undo annotations, modify the selection, etc.

This workflow transforms the problem from a \emph{recollection}
problem, where the user has to remember the correct alignment target
or otherwise spend time searching for it,
to a \emph{selection} problem, where the user can
pick the correct option from a list of candidates.
In the sTeX ecosystem, the snify tool has significantly increased
the efficiency of annotators, especially for novices who
are not deeply familiar with the domain model (i.e.\ the concept database).

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{img1.png}
    \caption{Screenshot of snify.}\label{fig:snify}
\end{figure}

In the background, snify uses a verbalization catalog
that links concept, here WikiData identifiers,
to their natural language renderings (verbalizations).
WikiData has both language-specific labels and synonyms.
For example, \lstinline.Q28920044. has the
label ``positive integer'' and lists ``natural number''
as a synonym.
This verbalization catalog drives the search for potential targets
and the selection of candidates.
To recognize inflected forms -- e.g.\ ``continuities'' instead of ``continuity''
-- snify uses an off-the-shelf stemmer.

As WikiData also has formula notations
for many concepts, including \LaTeX\ notations like \lstinline.\mathbb{N}.,
we can also use the same workflow to align formula fragments.
Such formula alignment is rather limited:
operators can be aligned, but their arguments are not marked, and
some operators, like the invisible multiplication in ``$ax$'',
are not represented by a glyph at all.
In sTeX, the solution is essentially to write down
a semantic representation of the formula and generate the presentation from it.
A grammar-based approach to generate the semantic representation
for ``conventional'' \LaTeX\ formulae
in collaboration with a human annotator was presented in\ednote{TODO: Luca's work}.
For alignments, this approach could be considered too invasive.

There is no reason to restrict the approach to \LaTeX\ documents.
As a case study, we have added support for HTML documents,
where the alignments are stored in HTML attributes.
Reading plain HTML in the command line interface is not ideal,
especially for complex MathML formulae.
To improve this, we have also created a browser-based interface
that closely imitates the command line interface
but can render HTML instead of showing the raw source (see \cref{fig:snify-browser}).

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{img2.png}
    \caption{Screenshot of the browser-based snify interface.
    Here, a MathML operator is being annotated.
    }\label{fig:snify-browser}
\end{figure}

In summary, we have demonstrated that the snify approach for sTeX annotation
can generalize to alignments with other concept databases
and document formats.
While the formula alignment is rather limited,
it was relatively easy to implement and could be improved in the future.
On a higher level, this shows how relatively simple and naive tools
can enable much more efficient alignment workflows,
and we hope to inspire the community to think about other tools
to help alignment authors -- be it text-to-database or database-to-database alignment.

% Document-to-database alignments are just annotations of specific
% text fragments or formula fragments in the document.
% For example, the text fragment ``natural number'' could be
% with the WikiData concept for natural numbers (Q21199).
% % For formulas, the annotation target can be less clear, for example because
% % some operations, such as the multiplication in ``$ax$'',
% % may not be represented by a glyph.
% % Depending on the document format, there there may still be
% % be an (invisible) symbol that can be annotated -- e.g.\ the unicode character for invisible multiplication.
% % An alternative is to annotate the whole sub-expression, i.e.\ the operation
% % including the operands.\ednote{these remarks break the flow}
% 
% Annotations can be created manually, e.g.\ by the authors,
% or automatically using machine learning techniques.
% Here, we focus on high-quality manual annotation,
% claiming that machine learning models are not sufficiently
% reliable, especially in research-level documents where training data is necessarily scarce.
% Presenting a user with wrong but convincing alignments
% is arguably worse than not providing any alignments at all.
% 
% The key challenge is to convice authors to actually create alignment annotations,
% as it is additional effort with little immediate benefit.
% To change the balance, we could either increase the benefit
% by providing better services based on the annotations,
% or reduce the effort by providing better annotation tools.
% Here we focus on the latter.
% 
% There is already an ecosystem of annotation tools, annotation formats
% and services based on them, developed in the KWARC group:
% \begin{compactitem}
% \item Annotation formats for \LaTeX (sTeX), HTML (FTML),
%     and, prototypically, Microsoft Word (WOIDE).
% \item Services for education (the ALeA platform)
% \item Tool support for manual annotation in sTeX
%     (snify and the sTeX IDE) and, prototypically, in Word (WOIDE).
% \end{compactitem}
% While there have been significant efforts to allow gradual adoption
% and remain as flexible as possible,
% potential users ultimately have to commit to a relatively
% complex ecosystem system of actively developed research software.
% As such, a dedicated teacher may be willing to invest time to
% move their teaching materials into this ecosystem
% to get the benefits of the ALeA learning platform,
% but a researcher writing a paper likely will not.
% 
% In this paper, we present a prototypical extension of the snify system
% for light-weight alignment independent of this ecosystem.
% The snify tool has been developed for the efficient annotation of natural language text
% with concept references and has been used to create an estimated 10,000 annotations.
% 
% Figure\ednote{TODO} shows
% 
% % Example:
% % Let $n < m$ be natural numbers.
% 
% 
% 
% % However, we can adapt some of the ideas and tools
% % for lightweight alignment of documents.
% % For example, the snify system has been designed
% % for large-scale manual creation of sTeX annotations.
% % By adapting it for alignment, we can
% % take advantage of
% 
% 
% % The focus of this paper is the alignment
% % of concept references in documents -- e.g.\ scientific papers -- with a concept database.
% % We will use WikiData as the concept database in our examples.
% % As WikiData is often referenced in other alignment databases,
% % e.g.\ MathGloss~\cite{mathgloss:arxiv},
% % this would also provide indirect alignments with other systems.
% % 
% % These document-to-database alignments are just annotations of specific
% % text fragments or formula fragments in the document.
% % For example, the text fragment ``natural number'' could be
% % with the WikiData concept for natural numbers (Q21199).
% % For formulas, the annotation target can be less clear, for example because
% % some operations, such as the multiplication in ``$ax$'',
% % are not represented by a glyph.
% % Depending on the document format, there there may still be
% % be an (invisible) symbol that can be annotated -- e.g.\ the unicode character for invisible multiplication.
% % An alternative is to annotate the whole sub-expression, i.e.\ the operation
% % including the operands.\ednote{these remarks break the flow}
% % 
% % Annotations can be created manually, e.g.\ by the authors,
% % or automatically using machine learning techniques.
% % Here, we focus on high-quality manual annotation,
% % claiming that machine learning models are not sufficiently
% % reliable, especially in research-level documents where training data is necessarily scarce.
% % Presenting a user with wrong but convincing alignments
% % is arguably worse than not providing any alignments at all.
% % 
% % The key challenge is to convice authors to actually create alignment annotations,
% % as it is additional effort with little immediate benefit.
% % To change the balance, we could either increase the benefit
% % by providing better services based on the annotations,
% % or reduce the effort by providing better annotation tools.
% % Here we focus on the latter.
% % 
% % There is already an ecosystem of annotation tools, annotation formats
% % and services based on them, developed in the KWARC group:
% % \begin{compactitem}
% % \item Annotation formats for \LaTeX (sTeX), HTML (FTML),
% %     and, prototypically, Microsoft Word (WOIDE).
% % \item Services for education (the ALeA platform)
% % \item Tool support for manual annotation in sTeX
% %     (snify and the sTeX IDE) and, prototypically, in Word (WOIDE).
% % \end{compactitem}
% % While there have been significant efforts to allow gradual adoption
% % and remain as flexible as possible,
% % potential users ultimately have to commit to a relatively
% % complex ecosystem system of actively developed research software.
% % As such, a dedicated teacher may be willing to invest time to
% % move their teaching materials into this ecosystem
% % to get the benefits of the ALeA learning platform,
% % but a researcher writing a paper will likely not.
% % 
% % However, we can adapt some of the ideas and tools
% % for lightweight alignment of documents.
% % For example, the snify system has been designed
% % for large-scale manual creation of sTeX annotations.
% % By adapting it for alignment, we can
% % take advantage of
%

\end{document}